---
title: "project code"
author: "Jiayi Song"
date: "2024-03-06"
output: html_document
---

## set up
```{r}
library(ggplot2)
library(maps)
library(dplyr)
library(tidyverse)
library(ggthemes)
```

```{r}
data = read.csv('storms.csv')
```

## 1. Damge from storms
*a)State Level Choropleth Maps*
```{r}
#modify data
data$total_damage <- data$DAMAGE_CROPS_USD + data$DAMAGE_PROPERTY_USD
names(data)[names(data) == "STATE"] <- "state"
damage_sum <- aggregate(total_damage ~ state, data = data, FUN = sum)
```

```{r}
#modify the map
us.states <- map_data("state") %>%
  as_tibble(.) %>%
  dplyr::rename(state = region) %>%
  select(-subregion) %>%
  mutate(state = toupper(state))

statenames <- as_tibble(
  cbind(state=state.name, state.abb = state.abb, 
        state.center.x = state.center$x, 
        state.center.y = state.center$y))
statenames <- statenames %>% mutate_each_(funs(as.numeric), 
                 vars=c("state.center.x","state.center.y"))
us.states <- left_join(us.states, statenames)
```

```{r}
#merge data
d_damage <- left_join(damage_sum, us.states, by='state')
```
```{r}
#generate the map
map_damage <- ggplot(d_damage, aes(x = long, y = lat, group = group, fill = total_damage)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(n.breaks = 5,low = "slategray2", high = "slategray4", name = "Total Damage") +
  theme_map() +
  labs(title = "Total Monetary Damage by State")+
  geom_text(data=statenames, inherit.aes = FALSE, 
            aes(label=state.abb, x=state.center.x, 
                y=state.center.y),size = 3, colour="gray25")
map_damage
```
<br>
*b) County Choropleth Maps*
```{r}
#modify data
names(data)[names(data) == "CZ_NAME"] <- "county"
damage_sum_c <- aggregate(total_damage ~ county + state, data = data, FUN = sum)
```

```{r}
#modify map
us.county <- map_data("county")%>%
  as_tibble(.) %>%
  dplyr::rename(state = region,county = subregion) %>%
  mutate(state = toupper(state),county = toupper(county))
```
```{r}
#merge data
d_damage_c <- left_join(damage_sum_c, us.county, by=c('state','county'))
```

```{r}
#generate the map
map_damage_c <- ggplot(d_damage_c, aes(x = long, y = lat, group = group, fill = total_damage)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(n.breaks = 5,low = "slategray2", high = "slategray4", name = "Total Damage") +
  theme_map() +
  labs(title = "Total Monetary Damage by County")
map_damage_c
```

<br>
*c) Density Map*
```{r}
#modify data
data$total_deaths <- data$DEATHS_DIRECT + data$DEATHS_INDIRECT
d_deaths <- subset(data, data$total_deaths > 0)
```


```{r}
map_0 <- map_data("state")
contour_deaths <- ggplot() + 
  geom_polygon(data = map_0, aes(x=long,y=lat,group=group), color = "grey", fill = NA) + theme_map() + 
  geom_density_2d(aes(x=BEGIN_LON,y=BEGIN_LAT, color="red"),data=d_deaths, size=1,bins = 10,na.rm = TRUE)
contour_deaths
```

```{r}
density_deaths <- contour_deaths +
  stat_density2d(data = d_deaths, geom = "polygon",bins = 10,na.rm = TRUE,
  aes(x = BEGIN_LON, y = BEGIN_LAT, fill = ..level..,alpha=..level..)) + 
  scale_fill_gradient(low = "yellow", high = "red") + theme_map() +
  labs(title = "Density Map of Severe Storms")+
  theme(legend.position = "right") #modify map
us.county <- map_data("county")%>%
as_tibble(.) %>%
dplyr::rename(state = region,county = subregion) %>%
mutate(state = toupper(state),county = toupper(county))
#merge data
d_damage_c <- left_join(damage_sum_c, us.county, by=c('state','county'))

density_deaths
```

<br>
Generally, I think the *density map* provides a better visual summary of the distribution of the destructive effects of storms. We can see that most of the states or counties suffer the same level of monetary damage in the Choropleth Maps, and it's difficult to distinguish them. However, the density map which presents the distribution of severe storms causing deaths, clearly shows the geographic differences.<br>
 
## 2. Location of Severe Events
*a) Interactive Map of Severe Weather Events* <br>
```{r}
library(leaflet)
```
```{r}
d_us <- subset(d_deaths, BEGIN_LON >= -125 & BEGIN_LON <= -65 & BEGIN_LAT >= 25 & BEGIN_LAT <= 50)
map_leaflet <- leaflet(d_us) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%  
  addMarkers(~BEGIN_LON, ~BEGIN_LAT, popup = paste("Event: ", d_us$EVENT_TYPE, "<br>",
                           "Number of deaths: ", d_us$total_deaths,"<br>",
                           "Year: ", d_us$YEAR,"<br>",
                           "End location: ",d_us$END_LOCATION))
map_leaflet
```


<br>
*b) Color by Type of Weather Event*
```{r}
table(d_us$EVENT_TYPE)
```
```{r}
d_us_new <- d_us
d_us_new$EVENT_TYPE <- ifelse(d_us_new$EVENT_TYPE %in% 
                c("Marine Strong Wind", "Marine Thunderstorm Wind","Waterspout"), "Marine Wind", d_us_new$EVENT_TYPE)
names(d_us_new)[names(d_us_new) == "BEGIN_LON"] <- "longitude"
names(d_us_new)[names(d_us_new) == "BEGIN_LAT"] <- "latitude"
```


```{r}
library(RColorBrewer)
pal = colorFactor("Spectral", domain = d_us_new$EVENT_TYPE) # Grab a palette
color_type = pal(d_us_new$EVENT_TYPE)
```
```{r}
map_leaflet_new<- leaflet(d_us_new) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%  
  addCircles(~longitude, ~latitude,color = color_type, radius = 50000,
             popup = paste("Event: ", d_us_new$EVENT_TYPE, "<br>",
                           "Number of deaths: ", d_us_new$total_deaths,"<br>",
                           "Year: ", d_us_new$YEAR,"<br>",
                           "End location: ",d_us_new$END_LOCATION)) %>%
  addLegend(pal = pal, values = ~d_us_new$EVENT_TYPE, title = "Event type")
map_leaflet_new
```


<br>
*c) Cluster*
```{r}
map_clust <- map_leaflet_new %>% addCircleMarkers(color = color_type, 
                       clusterOptions = markerClusterOptions())
map_clust
```








